---
title: "Reproducable Research Project 1"
author: "Frank Oelker"
date: "9 4 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Commit containing full submission
1. Code for reading in the dataset and/or processing the data  
2. Histogram of the total number of steps taken each day  
3. Mean and median number of steps taken each day  
4. Time series plot of the average number of steps taken  
5. The 5-minute interval that, on average, contains the maximum number of steps  
6. Code to describe and show a strategy for imputing missing data  
7. Histogram of the total number of steps taken each day after missing values are imputed  
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends  
9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report  

## Loading and preprocessing the data
The repository includes the needed File "activity.zip".  
This file will get unzipped into a CSV-file, if the CSV-file not exists.

```{r}
if(!file.exists("./activity.csv")) {
   dataSetZip <- "activity.zip"
   unzip(dataSetZip)
}
activity <- read.csv("activity.csv")
```

The dataset has to contain 17.568 observations.

```{r}
dim(activity)
```

Print some details of the dataset to check if a transformation is needed for the analysis.

```{r}
names(activity)
head(activity)
str(activity)
```


## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
